---
title: "HW2"
author: "Jason Gordon"
date: "2025-01-24"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidymodels)
```

# Homework 2 - EDA

Below is the work for Homework #2. I've kept the .csv files in my local environment to get them easily. I'll download them and do the preliminary checks.

```{r prelim}
airbnb <- read_csv("listings.csv")
codebook <- read_csv("codebook.csv")
airbnb |> dim() # 8269 rows, 75 cols.
```

Also going to check and remove NAs as needed. The columns `neighbourhood_group_cleansed` and `Calendar_updates` are all NA, so I'll remove them. However, the rows seem relatively full, so I'll keep all of them

```{r nas}
# Columns
na_cols <- apply(airbnb, 2, function(col){sum(is.na(col))})
na_cols |> sort(decreasing = TRUE) |> head(25)
missing_cols <- na_cols > (nrow(airbnb) * 0.5)
airbnb0 <- airbnb
airbnb <- airbnb[!missing_cols]

# Rows
na_rows <- apply(airbnb, 1, function(row){sum(is.na(row))})
na_rows |> sort(decreasing = TRUE) |> head(25)
```

## 3 - Cleaning Data

Now, let's clean the data and get a better idea of what's going on.

```{r check1}
tibble(airbnb)
```

With this data, I'll be transforming the following 10 variables:

* host_location (factorize by state/country, split city)
* host_response_time (factorize)
* host_response_rate(% -> int)
* host_acceptance_rate (% -> int)
* host_neighbourhood (factorize)
* host_verifications (This is a split list of 3, so doing 3 True/False columns.)
* neighbourhood_cleansed (factorize)
* bathrooms_text (factorize)
* price ($ -> int)
* reviews_per_month(filter)

### Creating Numbers

For `host_response_rate`, `host_acceptance_rate`, and `price`, we'll need to create integers, which I'll do accordingly.

```{r response_rate}
response_rate <- airbnb$host_response_rate |>
  gsub("%", "", x = _) |>
  as.numeric() # coerces NAs automatically
hist(response_rate,
     breaks = 100,
     main = "Host Response Rate",
     xlab = "Response Rate (%)") # Making sure there aren't any outside of 0 -> 100.
# But...why are there some so low?
sum(response_rate < 10, na.rm = TRUE) # Hmm...I'll filter these out.
airbnb$host_response_rate <- response_rate
airbnb <- airbnb |> filter(response_rate >= 10)
```
```{r acceptance_rate}
acceptance_rate <- airbnb$host_acceptance_rate |>
  gsub("%", "", x = _) |>
  as.numeric() # coerces NAs automatically
hist(acceptance_rate,
     breaks = 100,
     main = "Host Acceptance Rates",
     xlab = "Aceptance Rate (%)") # 0% acceptance??
sum(acceptance_rate < 10, na.rm = TRUE) # Quite a few with low acceptance. Will filter.
airbnb$host_acceptance_rate <- acceptance_rate
airbnb <- airbnb |> filter(host_acceptance_rate >= 10)
hist(airbnb$host_acceptance_rate,
     breaks = 100,
     main = "Host Acceptance Rates (Fixed)",
     xlab = "Aceptance Rate (%)")
```
```{r price}
price <- airbnb$price |>
  gsub("\\$|,", "", x = _) |>
  as.numeric()
hist(log(price),
     breaks = 100,
     main = "Distribution of Airbnb Prices (Histogram)",
     xlab = "log(price)") # Hmm...why are some over 9?
boxplot(log(price), horizontal = TRUE) # Ok, so it's not *that* much of an outlier.
title("Distribution of Airbnb Prices", xlab = "log(price)")
sort(price, decreasing = TRUE) |> head(25)
# I mean, ok, I guess...$10k/night isn't bad if it's 14 rooms like I think I saw.
airbnb$price <- price
```

In these, I've made a few histograms to see if there are any major outliers. If there was any specific issues, I also used a boxplot to find how bad outliers were.

We'll also filter the amount of reviews per month to remove the massive outliers.

```{r reviewspermonth}
boxplot(airbnb$reviews_per_month, horizontal = TRUE) # 60 reviews per month seems like an outlier...
title("Distribution of Reviews per month", xlab = "Reviews/month")
sort(airbnb$reviews_per_month, decreasing = TRUE) |> head(25)
# Let's filter out more than 20. That's strange for it being so high.
airbnb <- airbnb |> filter(reviews_per_month < 20)
boxplot(airbnb$reviews_per_month, horizontal = TRUE) # Much better.
title("Distribution of Reviews per Month (Fixed)", xlab = "Reviews/month")
```

### Modifying and Factorizing Text

Now, I'll factorize categorical variables, making them easier to use later.

```{r response_time}
airbnb |> group_by(host_response_time) |> summarize(count = n())
response_time <- airbnb$host_response_time |>
  gsub("N/A", NA, x = _) |>
  factor()
head(response_time) # Pretty clean.
airbnb$host_response_time <- response_time
```
```{r neighbourhood}
airbnb |> group_by(host_neighbourhood) |> summarize(count = n())
neighborhood <- airbnb$host_neighbourhood |>
  factor()
head(neighborhood)
airbnb$host_neighbourhood <- neighborhood
```
```{r cleansed}
airbnb |> group_by(neighbourhood_cleansed) |> summarize(count = n())
neighborhood_cleansed <- airbnb$neighbourhood_cleansed |>
  factor()
head(neighborhood_cleansed)
airbnb$neighbourhood_cleansed <- neighborhood_cleansed
```
```{r textbath}
airbnb |> group_by(bathrooms_text) |> summarize(count = n())
bath_text <- airbnb$bathrooms_text |>
  factor()
head(bath_text)
airbnb$bathrooms_text <- bath_text
```

### Other Modifications

First, I'll split city and state/country on the host location, factorizing one but keeping the other as a string

```{r location}
airbnb |> group_by(host_location) |> summarize(count = n())
# Let's split these since they all follow [city], [state/country].
loc <- airbnb$host_location |> str_split(", ")
head(loc)
# Hmm...there seems to be just "Italy" there. Let's select the *last* element of the list.
state <- sapply(loc, tail, 1)
city <- sapply(loc, function(x){ifelse(length(x) == 1, NA, x[1])})
# City can be a string. I'll make state/country a factor.
state <- state |> factor()
airbnb$host_city <- city
airbnb$host_state_country <- state
```

`host_verification` is also a list of three categories, so it's easier to split them up to do True/False for each one.

```{r host_verification}
airbnb |> group_by(host_verifications) |> summarize(n())
verif <- airbnb$host_verifications
head(verif) # Needs cleaning first.
verif <- verif |>
  gsub("\\[|\\]|'", "", x = _) |>
  str_split(", ")
email <- sapply(verif, function(x){ifelse("email" %in% x, TRUE, FALSE)})
# Verification
sum(email)
sum(!email)
# Now for the other two.
phone <- sapply(verif, function(x){ifelse("phone" %in% x, TRUE, FALSE)})
work_email <- sapply(verif, function(x){ifelse("work_email" %in% x, TRUE, FALSE)})
airbnb$email <- email
airbnb$phone <- phone
airbnb$work_email <- work_email
head(select(airbnb, email, phone, work_email)) # Let's see if we did it right.
```

## Amenities

For `amenities`, we'll first break the entire column into a list of lists, and then use one of the lists to create a True/False dataframe of the rest.

I used the 5th term of `amenities` here since it was relatively long with 59 elements and many of the terms were rather simple, making it easy for me to use te `%in%` function to find terms.

```{r amenities}
head(airbnb$amenities) # Ok, so there's a lot going on here.
# Let's strip the brackets and begin breaking it down.
amenities <- airbnb$amenities |>
  gsub("\\[\"|\"\\]", "", x = _) |> # Strips beginning and end.
  str_split("\\\", \\\"") # Should be \", \"
head(amenities)

names <- amenities[[5]] # Most Extensive + Simplest
df <- sapply(names, function(term){
  sapply(amenities, function(x){ifelse(term %in% x, TRUE, FALSE)})
}) |> data.frame()

# Seeing how many are just all False
false_cols <- apply(df, 2, function(col){sum(col)})
false_cols |> sort() |> head(25)
hist(false_cols)

# I'll cut anything with under 1/10th.
empty_cols <- false_cols > (nrow(df) * 0.9)
df <- df[!empty_cols]

airbnb <- cbind(airbnb, df)
```

## Transformations

Overall, I think the price is useful as-is (with it becoming an int). However, having general statistics for the neighborhood would be useful.

```{r transformation}
head(airbnb$price)
# I think this is pretty useful as-is (with the commas and $ stripped.)
# If we're trying to predict price ranges based on the market, maybe stats as
# mentioned in #7 could be useful
stat_df <- airbnb |>
  filter(host_city == "Chicago") |>
  group_by(host_neighbourhood) |>
  summarize(count = n(),
            mean_price = mean(price, na.rm = TRUE),
            median_price = median(price, na.rm = TRUE),
            hosts = length(unique(host_id)),
            avail_30 = sum(ifelse(availability_30, 1, 0)),
            pct_avail = (avail_30 / count) * 100)
airbnb <- airbnb |> merge(stat_df, by = "host_neighbourhood")
```

## Questions (#7)

For these questions, I arranged the dataframe according to the mean/median or count/percentage to easily show some basic statistics as asked.

I could also call head(1) for these, but I feel it's more insightful to have the entire tibble.

```{r statq}
arrange(stat_df, desc(mean_price)) # Highest Price
arrange(stat_df, desc(median_price))
arrange(stat_df, mean_price) # Lowest Price
arrange(stat_df, median_price)
arrange(stat_df, desc(hosts)) # Hosts
arrange(stat_df, desc(avail_30)) # Availability
arrange(stat_df, desc(pct_avail))
```